Hierarchical Inference of the Intrinsic Ellipticity Distribution of Galaxies during Weak Lensing Analysis

### 2013-03-14 docs/ms.tex
Update ms.tex

\subsection{Likelihood Functions}
If we know the ellipticity distribution of unlensed galaxies based
on some parasmeters $\alpha$,the likelihood of reduced shear $g$
and parameter $\alpha$ is like:

@@

Note that both $\epsilon_{0}$ and $\epsilon_{\ell}$are complex numbers.So
$|\frac{\partial\epsilon_{0}}{\partial\epsilon_{\ell}}|$ is the Jacobian
determinant term.Here the prior distribtuion we use for $\vec{g}$
is a 2Dflat distribution in the region $|g|<1$.Therefore $prior(\vec{g)}=\begin{cases}
1 & |\vec{g}|<1\\
0 & |\vec{g}|>1
\end{cases}$

@@

\subsection{model for P($\epsilon_{0}$)}
Two flexible models are tested for P($\epsilon_{0}$)
One is Beta function model and one is step function model.
While using step function model we use a prior regularzation on $\alpha$ 
\begin{equation}
Prior(\vec{\alpha}|\xi)=exp(-\xi\sum(\alpha_{i}-\alpha_{i-1})^{2})
\end{equation}

\subsection{MCMC Sampler}

@@

>>>>

\subsection{Likelihood Function}

We assume Lensed galaxy ellipticity $\epsilon_{\ell}$ is related
to the intrinsic galaxy ellipticity $\epsilon_{0}$ in the weak lensing
regime via
\begin{equation}
\epsilon_{\ell}=\frac{\epsilon_{0}+g}{1+g^{*}\epsilon_{0}}
\end{equation}

and it's inverse formula 
\begin{equation}
\epsilon_{0}=\frac{\epsilon_{\ell}-g}{1-g^{*}\epsilon_{\ell}}
\end{equation}
from Schramm \& Kayser (1995), Seitz \& Schneider (1997), where $\epsilon$
is represented as a complex variable and g, $g^{*}$ are the reduced
shear and its complex conjugate, respectively. $\epsilon$ is defined
in terms of the major and minor axes and orientation a, b, $\phi$,
respectively, as e = (a − b)/(a + b) exp(2i$\phi$). Due to the isotropic
of Universe ,we expect $\phi$has a uniform distribution from 0 to
$\pi$.Then we can expect 
\begin{equation}
<\epsilon_{\ell}>=g
\end{equation}

So $<\epsilon_{\ell}>$ can be adopted as an estimator of g.

However in our hierachical Bayesian approach we are able to obtain
the posterior distribution of g ,starting from the same formulas.

Suppose we know the ellipticity distribution of unlensed galaxies
based on some parasmeters $\alpha$,the likelihood of reduced shear
$g$ and parameter $\alpha$ is like:

@@

Note that both $\epsilon_{0}$ and $\epsilon_{\ell}$are complex numbers.So
$|\frac{\partial\epsilon_{0}}{\partial\epsilon_{\ell}}|$ is the Jacobian
determinant term,which can be calculated numerically from formula
(5).

If there is measurement error on $\epsilon_{ob}$,we will have 
\begin{equation}
L(\vec{g},\alpha|\epsilon_{ob})=p(\epsilon_{ob}|\vec{g},\alpha)=\int f(\epsilon_{ob}|\epsilon_{\ell})p(\epsilon_{\ell}|\vec{g},\alpha)
\end{equation}
Here the prior distribtuion we use for $\vec{g}$ is a 2Dflat distribution
in the region $|g|<1$.Therefore $prior(\vec{g)}=\begin{cases}
1 & |\vec{g}|<1\\
0 & |\vec{g}|>1
\end{cases}$

@@

One thing to note is that the likelihood of g in patch i only depends
on the shape parameters but not dependent to g in other patch.Therefore,the
marginalized likelihood of shape parameters has an equation like:

\begin{equation}
\ell(\alpha|data)=\prod_{i}\ell_{i}(\alpha|data_{i})
\end{equation}

$\ell_{i}(\alpha|data_{i})$ is the marginalized likelihood function
calculated with data in the ith patch.From this formula we can see
how better estimation of $\alpha$ will be obtained if we infer more
sky patches simultaneously.

\subsection{Model for Prior($\epsilon_{0}$)}

The correctness of posterior distribution of g depends on the intrinsic
ellipticity distribution P($\epsilon_{0}|\alpha$). Due to the limitation
of human's knowledge we are not able to give an analytical form of
$P(\epsilon_{0}|\alpha)$.Therefore we proposed an exceedingly flexible
model for $P(\epsilon_{0}|\alpha)$,which does not requires further
assumption on $P(\epsilon_{0}|\alpha)$.

We divide the 2-D space r$\in$(0,1) into N bins .The probability
of each bin is $p_{i}=e^{\alpha_{i}}$ .$\overrightarrow{\alpha}$
is a N-D parameter vector to be infered,simultaneously with g.

From practice we find that a prior smooth regularzation on $\alpha$
is needed to obtain a good estimation on shear g.

The prior distribution we used on $\alpha$ is : 
\begin{equation}
Prior(\vec{\alpha}|\xi)=exp(-\xi\sum(\alpha_{i}-\alpha_{i-1})^{2})
\end{equation}

Here $\xi$ is a prefactor need to be tuned .

Meanwhile,in order to show the correctness of our method ,we also
set up a simulation in which we know the true $P(|\epsilon_{0}||\alpha)$
is a Beta distribution ,and thus $\alpha$ is the 2-D parameter (p,q)
of the Beta distribution.We will show the posterior samples using
both of these two methods in next section. Note that for either model
we assume P($\epsilon_{0}$) is isotropic.

\subsection{Gibbs Sampler}

@@

In Gibbs sampling ,samples of variable $x_{j}$ is made from the conditional
probability distribution:$p(x_{j}^{(i)}|x_{1}^{(i)},....,x_{j-1}^{(i)},x_{j+1}^{(i-1)}...,x_{n}^{(i-1)})$

While the conditional distribution of one variable given all others
is just proportional to the joint distribution

\begin{equation}
p(x_{j}|x_{1},....,x_{j-1},x_{j+1}...,x_{n})\varpropto p(x_{1},.....,x_{n})
\end{equation}

One thing to note is that the conditional probability of g in ith
patch only explicitly depends on $\alpha$ ,but not g in other patches.Therefore
with $\alpha$ fixed ,the posterior distribution of reduced shear
in ith patch are independent from each other .
\begin{equation}
p(\vec{g}_{1}....\vec{g}_{n}|\alpha,Data)=p(\vec{g}_{1}|\alpha,Data_{1})....p(\vec{g}_{n}|\alpha,Data_{n})
\end{equation}
Therefore sampling reduced shear g from conditional distribution can
be reduced to 2D sampling problem,which can be done very quickly.

The total time needed to sample $g$ and $\alpha$ of 64 sky patches
with 8 galaxies in each patch will only be approximitly 25s on a 2
GHz CPU.

\subsection{Simulations}

The procedure of our simulation is that ,first we generated N{*}NP
unlensed galaxies' elliprticity from the ``true'' $P(\epsilon_{0})$,and
asign these galaxies into NP patches equally.So each patch has N galaxies.Second
we generate NP true reduced shears ,and asign them to each patch,and
then shear the unlensed galaxies using formula (4) .

Finally we can put on some noise on $\epsilon_{\ell}$ to get the
observed ellipticity $\epsilon_{ob}$. For simplicity we've only considered
the situation when P($\epsilon_{ob}|\epsilon_{\ell})$ is a 2D Gaussian.

Use Gibbs sampler and $P(\epsilon_{0}|\alpha)$ mentioned previously,we
are to sample the posterior distribution of g and $\alpha$ simultaneously
and effectively.
